\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-01-20 redaction/dhp_sept2016.tex
chnagement section 2

We estimate international transport costs at the 3-digit classification level, even if data series on the cif and fas prices are available at the 5-digit classification level. As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time. Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013. Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation. We thus obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels. Estimation results at the 4-digit classification level are reported in Appendix ??.} Depending on the considered year, this leaves us with around 200 sectors, from approximately 200 countries of origin.

@@

\begin{equation*}
\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
\end{equation*}

\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2)} this separability is in a multiplicative way for the former and an additive way for the latter.In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k \simeq (\tau_i-1)\times (\tau_k -1)-1$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.}

@@

The non-linearity of Equation (\ref{eq:equation0}) implies that it cannot be estimated using standard linear estimators. All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations. The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next. See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time. Confronted to this arbitrage, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$). Put it differently, this amounts making the additional assumption, that all products $k$ in a 3-digit sector $s$ share the
same structure of costs.\footnote{We ensure the robustness of our estimation results to this assumption in two ways. First, we conduct the estimations at the 4-digit level for some selected years. As reported in Appendix \ref{app:4digit}, the results are not substantially affected. Second, we provide a decomposition variance exercise on the observed cif-fas price. As reported in Appendix ???, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.} This drives us to estimate a modified version of Equation (\ref{eq:equation0}), according to:
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
\end{equation}
where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0. To eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables. These cut-offs are aimed at eliminating reporting or coding errors. We estimate this equation for each year over the period 1974-2013, for each of the transportation mode reported (air or vessel), on a sectoral-origin country basis ($i,s$). Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode (at the 3-digit level).   \smallskip

@@

\begin{eqnarray}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i}\times\tau_{s(k)}-1 \right) + \epsilon^{nlI}_{ik} \label{eq:model_nlI} \\
%\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik} \label{eq:model_nlA} \\
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{s(k)}-1 +\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_IetA}
\end{eqnarray}

After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{is(k)} = \widehat{\tau_{i}} \times \widehat{\tau_{s(k)}}$ and $\widehat{t}^{add}_{is(k)} = \widehat{t}_{i} + \widehat{t}_{s(k)}$, that is country-sector specific, by year and transport mode. When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{nlI}_{is(k)} = \widehat{\tau_{i}} \times \widehat{\tau_{s(k)}}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form. To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}$^{,}$\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs. This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik}$. The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results. For sake of space saving, the results of this model are not reported in the paper. Of course, they are available upon request to the authors.} Similarly as \citet{Irrazabal_2015}, we take the average over the product-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme. We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{nlI}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode. These results are reported in Section \ref{sec:results_decomposition}.

@@

The objective of this section is twofold. First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is included or not in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})). Second, we evaluate the importance of the additive component of in overall transport costs through the means of goodness-of-fit measures.

@@

\subsection{Fab-cif prices}

@@

\section{More on the estimation results \label{app:more_results}}

In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})).

\subsection{Detailed results at the 3-digit classification level}

In this section, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2013, when the degree of classification retained ($s$) is at the 3-digit classification level. Table \ref{tab:result_air_3d_detail} reports the results for Air transport. The results for Ocean transport are displayed in Table \ref{tab:result_ves_3d_detail}.

@@

\begin{table}[htbp]
  \centering
  \caption{Vessel: Dynamics of Trade Costs over Time (3 digits), selected years}
\begin{center}
    \begin{tabular}{l|cccccc}
   \hline\hline
Year         & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\
 \hline
   \multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs} }  \\
Mean  & 1.098 & 1.065 & 1.057 & 1.051 & 1.040 & 1.036  \\
Median & 1.096 & 1.055 & 1.046 & 1.049 & 1.036 & 1.033  \\
Standard Error & 0.053 & 0.040 & 0.032 & 0.028 & 0.020 & 0.018  \\
\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs} }    \\
\hline
\multicolumn{7}{l}{\textit{Additive term } }  \\ \hline
Mean  & 0.051 & 0.034 & 0.027 & 0.028 & 0.025 & 0.015  \\
Median & 0.029 & 0.023 & 0.017 & 0.022 & 0.019 & 0.008 \\
Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\ \hline
\multicolumn{7}{l}{\textit{Iceberg term} } \\ \hline
Mean  & 1.054 & 1.031 & 1.033 & 1.025 & 1.019 & 1.022 \\
Median & 1.049 & 1.024 & 1.028 & 1.021 & 1.018 & 1.018  \\
Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018  \\
\hline
 \# obs & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 \\ \hline
\hline\hline
\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fas price.}}}
\end{tabular}%
\end{center} \label{tab:result_ves_3d_detail}
\end{table}%

\subsection{Estimation at the 4-digit level \label{app:4digit}}

\paragraph{Transport cost estimates} Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in Air and Ocean transport respectively.

\paragraph{Goodness-of-fit tests at the 4-digit level} We now report the goodness-of-fit exercise (conducted by transport mode) at the 4-digit product classification level (for the selected years). The results are reported in Tables \ref{tab:good_fit_air_rob} (for Air) and \ref{tab:good_fit_ves_rob} (for Vessel).
\begin{table}[htbp]
  \centering
  \caption{Air: Measures of Goodness-of-fit, 4-digits}
\begin{center}
\label{tab:good_fit_air_rob}%
%\vspace{0.5cm}
%\scalebox{0.97}{
\begin{tabular}{l|cccccc}
\hline
\hline
      & \multicolumn{6}{c}{Year}              \\
      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\
\hline
\textbf{R2} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &       &      \\
Term I only & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
Terms A \& I & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\
\textbf{SER} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{}  \\
Term I only & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
Terms A \& I & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
Term I only & \multicolumn{1}{c}{-17505.55} & \multicolumn{1}{c}{-21813.46} & \multicolumn{1}{c}{-30960.56} & \multicolumn{1}{c}{-44067.62} & \multicolumn{1}{c}{-49375.57} & \multicolumn{1}{c}{-53197.87}  \\
Terms A\& I & \multicolumn{1}{c}{-14895.81} & \multicolumn{1}{c}{-18589.91} & \multicolumn{1}{c}{-26553.53} & \multicolumn{1}{c}{-37297.93} & \multicolumn{1}{c}{-45747.57} & \multicolumn{1}{c}{-49899.14}  \\
\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{}  \\
Term I only & \multicolumn{1}{c}{36243.10} & \multicolumn{1}{c}{44966.91} & \multicolumn{1}{c}{63417.12} & \multicolumn{1}{c}{89747.24} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.73} \\
Terms A \& I & \multicolumn{1}{c}{31873.63} & \multicolumn{1}{c}{39495.82} & \multicolumn{1}{c}{55777.05} & \multicolumn{1}{c}{77439.85} & \multicolumn{1}{c}{94059.14} & \multicolumn{1}{c}{102224.28}  \\
\textbf{Test LL} &       &       &       &       &       &       \\
2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.47} & \multicolumn{1}{c}{6447.09} & \multicolumn{1}{c}{8814.06} & \multicolumn{1}{c}{13539.39} & \multicolumn{1}{c}{7255.99} & \multicolumn{1}{c}{6597.45}  \\
\# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818}  \\
p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
\hline\hline
\multicolumn{7}{l}{\parbox[l]{15cm}{ \vspace{7pt}\scriptsize{Notes: R$^{2}$ between the log of predicted ratio and the log of the observed ratio. The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}}}
\end{tabular}%
\end{center}
\end{table}%

@@

>>>>

We estimate international transport costs at the 3-digit classification level, even if data series on the cif and fas prices are available at the 5-digit classification level. As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time. Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013. Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation. We thus obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels. Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the considered year, this leaves us with around 200 sectors, from approximately 200 countries of origin.

@@

\begin{equation}
\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
\end{equation}

\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2)} this separability is in a multiplicative way for the former and an additive way for the latter. In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form does not make a substantial difference since, for small values (as we obtain), we have $\tau_i\times \tau_k \simeq (\tau_i-1)\times (\tau_k -1)-1$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.}

@@

The non-linearity of Equation (\ref{eq:equation0}) implies that it cannot be estimated using standard linear estimators. All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations. The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next. See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time. Confronted to this arbitrage, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$). Put it differently, this amounts making the additional assumption, that all products $k$ in a 3-digit sector $s$ share the
same structure of costs.\footnote{We ensure the robustness of our estimation results to this assumption in two ways. First, we conduct the estimations at the 4-digit level for some selected years. As reported in Appendix \ref{app:4digit}, the results are not substantially affected. Second, we provide a decomposition variance exercise on the observed cif-fas price. As reported in Appendix ???, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.} This drives us to estimate a modified version of Equation (\ref{eq:equation0}), specified as:
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
\end{equation}
where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0. To eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables. These cut-offs are aimed at eliminating reporting or coding errors. We estimate this equation for each year over the period 1974-2013, for each of the transportation mode reported (air or vessel), on a sectoral-origin country basis ($i,s$). Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode (at the 3-digit level).   \smallskip

@@

\begin{eqnarray}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i}\times\tau_{s(k)}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI} \\
%\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon^{nlA}_{ik} \label{eq:model_nlA} \\
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{s(k)}-1 +\frac{t_{i} + t_{s}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_IetA}
\end{eqnarray}

After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$, that is country-sector specific, by year and transport mode. When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form. To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.}$^{,}$\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs. This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$. The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results. For sake of space saving, the results of this model are not reported in the paper. Of course, they are available upon request to the authors.} Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme. We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{ice}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode. These results are reported in Section \ref{sec:results_decomposition}.

@@

The objective of this section is twofold. First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})). Second, we evaluate the importance of the additive component of in overall transport costs through the means of goodness-of-fit measures.

@@

@@

\section{Estimation at the 3-digit classification level \label{app:more_results}}

\subsection{Transport costs estimates: More detailed results}

In this section, we report more detailed results for the estimates for international transport costs, by transport mode on a yearly basis, when either additive costs are included in the estimation (Equation (\ref{eq:model_IetA})) or not (Equation (\ref{eq:model_nlI})), under our benchmark sectoral classification level (3 digit). Precisely, we complement the results displayed in Table \ref{tab:summary_results} by reporting the estimates of international transport costs for a sample of years over 1974-2013, when the degree of classification retained ($s$) is at the 3-digit classification level. Table \ref{tab:result_air_3d_detail} reports the results for Air transport. The results for Ocean transport are displayed in Table \ref{tab:result_ves_3d_detail}.

@@

\begin{table}[htbp]
  \centering
  \caption{Vessel: Transport costs estimates, 3 digit (selected years)}
\begin{center}
    \begin{tabular}{l|cccccc}
   \hline\hline
Year         & 1974  & 1980  & 1990  & 2000  & 2010  & 2013   \\
 \hline
   \multicolumn{7}{l}{\textbf{With only Iceberg Trade Costs} }  \\
Mean  & 1.098 & 1.065 & 1.057 & 1.051 & 1.040 & 1.036  \\
Median & 1.096 & 1.055 & 1.046 & 1.049 & 1.036 & 1.033  \\
Standard Error & 0.053 & 0.040 & 0.032 & 0.028 & 0.020 & 0.018  \\
\multicolumn{7}{l}{\textbf{With Additive \& Iceberg Trade Costs} }    \\
\hline
\multicolumn{7}{l}{\textit{Additive term } }  \\ \hline
Mean  & 0.051 & 0.034 & 0.027 & 0.028 & 0.025 & 0.015  \\
Median & 0.029 & 0.023 & 0.017 & 0.022 & 0.019 & 0.008 \\
Standard Error & 0.085 & 0.046 & 0.040 & 0.043 & 0.025 & 0.020 \\ \hline
\multicolumn{7}{l}{\textit{Iceberg term} } \\ \hline
Mean  & 1.054 & 1.031 & 1.033 & 1.025 & 1.019 & 1.022 \\
Median & 1.049 & 1.024 & 1.028 & 1.021 & 1.018 & 1.018  \\
Standard Error & 0.041 & 0.023 & 0.022 & 0.021 & 0.018 & 0.018  \\
\hline
 \# obs & 19007 & 17356 & 28383 & 36090 & 37748 & 38473 \\ \hline
\hline\hline
\multicolumn{7}{l}{\parbox[l]{11cm}{ \vspace{7pt}\scriptsize{Notes: Statistics are obtained weighting each observation by its value in transport (mode-dependent). Additive term expressed in fraction of fas price.}}}
\end{tabular}%
\end{center} \label{tab:result_ves_3d_detail}
\end{table}%

@@

\section{Estimation at the 4-digit level \label{app:4digit}}

\subsection{Transport cost estimates}

Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in Air and Ocean transport respectively.

\subsection{Goodness-of-fit tests at the 4-digit level}

We now report the goodness-of-fit exercise (conducted by transport mode) at the 4-digit product classification level (for the selected years). The results are reported in Tables \ref{tab:good_fit_air_rob} (for Air) and \ref{tab:good_fit_ves_rob} (for Vessel).
\begin{table}[htbp]
  \centering
  \caption{Air: Measures of Goodness-of-fit, 4-digits}
\begin{center}
\label{tab:good_fit_air_rob}%
%\vspace{0.5cm}
%\scalebox{0.97}{
\begin{tabular}{l|cccccc}
\hline
\hline
      & \multicolumn{6}{c}{Year}              \\
      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\
\hline
\textbf{R2} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       &       &      \\
Term I only & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.49} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.50} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.35} \\
Terms A \& I & \multicolumn{1}{c}{0.63} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.65} & \multicolumn{1}{c}{0.66} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{0.45} \\
\textbf{SER} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{}  \\
Term I only & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.88} & \multicolumn{1}{c}{0.93}  \\
Terms A \& I & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{0.80} & \multicolumn{1}{c}{0.86}  \\
\textbf{Log-likelihood} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
Term I only & \multicolumn{1}{c}{-17505.55} & \multicolumn{1}{c}{-21813.46} & \multicolumn{1}{c}{-30960.56} & \multicolumn{1}{c}{-44067.62} & \multicolumn{1}{c}{-49375.57} & \multicolumn{1}{c}{-53197.87}  \\
Terms A\& I & \multicolumn{1}{c}{-14895.81} & \multicolumn{1}{c}{-18589.91} & \multicolumn{1}{c}{-26553.53} & \multicolumn{1}{c}{-37297.93} & \multicolumn{1}{c}{-45747.57} & \multicolumn{1}{c}{-49899.14}  \\
\textbf{AIC criteria} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &       & \multicolumn{1}{c}{}  \\
Term I only & \multicolumn{1}{c}{36243.10} & \multicolumn{1}{c}{44966.91} & \multicolumn{1}{c}{63417.12} & \multicolumn{1}{c}{89747.24} & \multicolumn{1}{c}{100317.13} & \multicolumn{1}{c}{107963.73} \\
Terms A \& I & \multicolumn{1}{c}{31873.63} & \multicolumn{1}{c}{39495.82} & \multicolumn{1}{c}{55777.05} & \multicolumn{1}{c}{77439.85} & \multicolumn{1}{c}{94059.14} & \multicolumn{1}{c}{102224.28}  \\
\textbf{Test LL} &       &       &       &       &       &       \\
2$\times$(ll(UR) -ll(R)) & \multicolumn{1}{c}{5219.47} & \multicolumn{1}{c}{6447.09} & \multicolumn{1}{c}{8814.06} & \multicolumn{1}{c}{13539.39} & \multicolumn{1}{c}{7255.99} & \multicolumn{1}{c}{6597.45}  \\
\# restrictions  & \multicolumn{1}{c}{640} & \multicolumn{1}{c}{698} & \multicolumn{1}{c}{778} & \multicolumn{1}{c}{833} & \multicolumn{1}{c}{824} & \multicolumn{1}{c}{818}  \\
p-value & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
\hline\hline
\multicolumn{7}{l}{\parbox[l]{15cm}{ \vspace{7pt}\scriptsize{Notes: R$^{2}$ between the log of predicted ratio and the log of the observed ratio. The number \# of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products.}}}
\end{tabular}%
\end{center}
\end{table}%
